Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2020-08-26 mice_synthesizing/paper.Rmd
Link synthetic and missing data.

---
title: "Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*"
subtitle: "Anony*mice*d shareable data: Using *mice* to create multiply imputed synthetic datasets" 
author:
- Thom Volker, Gerko Vink & Stef van Buuren
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
bibliography: federated_imp.bib
csl: "/Users/thomvolker/Documents/styles/apa-6th-edition.csl"
---

@@

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general concept was to release multiply imputed synthetic datasets, instead of the actually observed data. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available datasets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, say, income values above a certain threshold. Additionally, a researcher can opt to simply replace all values of every observation.

Potentially, this approach could fulfil the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data.

>>>>

---
title: "Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*"
subtitle: "Anony*mice*d shareable data: Using *mice* to create multiply imputed synthetic datasets" 
author:
- Thom Volker, Gerko Vink & Stef van Buuren
date: "`r format(Sys.time(), '%d-%m-%Y')`"
output: 
  html_document:
    theme: spacelab
    highlight: tango
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
bibliography: federated_imp.bib
csl: "apa-6th-edition.csl"
---

@@

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general concept was to release multiply imputed synthetic datasets, instead of the actually observed data. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed dataset as a whole with multiple synthetic versions. Additionally, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available datasets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, say, income values above a certain threshold.

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation for missing data, as proposed by @rubin_multiple_1987. Now, instead of replacing the missing values with multiple draws from the posterior predictive distribution, the sensitive values are replaced accordingly. Similarly to analysing multiply imputed datasets dealing with missing data, multiply imputed synthetic datasets can be analysed using conventional analysing techniques. The results of the analyses on each synthetic dataset should be pooled, so that the researcher obtains the estimates of the quantity of interest, as well as the estimate of its variance that adequately account for the additional variability that is induced by the imputation procedure.

Potentially, this approach could fulfil the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data are highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these datasets should be kept at a minimum.

Since multiple imputation of synthetic data is closely linked to multiple imputation for missing data, using existing missing data software that research are familiar with already, would dramatically reduce the effort of creating synthetic data. The R-package `mice` [@mice] is particularly suited for generating synthetic data, since it already implements multiple imputation for missing data in a straightforward and user-friendly manner. However, `mice` is not restricted to imputation for missing data, since it can be easily adjusted to impute every observed value.

***IK DENK DAT HET GOED IS OM HIER NOG TWEE KORTE ZINNEN OVER MICE NEER TE ZETTEN, MAAR IK VERWACHT DAT JULLIE BETER WETEN HOE EN WAT***

To fully utilize the potential gains of synthetic data, while keeping the costs

To fully

Thus, the potential gains of using synthetic data are very high, if the costs of creating it can be kept at a minimum.

If the approach to creating synthetic datasets can be simplified

<!--

It is easy to draw
